Supplementary information

### 2018-06-21 manuscript/02_methods.Rmd
Add some correlation text to manuscript

The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework (OSF) at <https://osf.io/w8y73/>.
The TRY data request used for this analysis has been archived at <http://try-db.org>,
and can be retrieved by providing the request ID (#1584).
Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (ashiklom@bu.edu).

>>>>

The magnitude of correlation coefficients can be strongly influenced by data sampling, particularly sample size and range (correlations are likely to be stronger when data have more samples and a larger range).
To evaluate the contribution of uneven data sampling to our correlation predictions, we plotted the square of each pairwise correlation coefficient against pairwise sample size and normalized data range (Fig.\ S1 and S2).

The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework at <https://osf.io/w8y73/>.
The TRY data request used for this analysis has been archived at <http://try-db.org>,
and can be retrieved by providing the TRY data request ID (#1584).
Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (ashiklom@bu.edu).

### 2018-06-21 manuscript/03_results.Rmd
Fix manuscript typos. Add cover letter.
Add some correlation text to manuscript

For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.\ 4).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs, tropical broadleaved deciduous trees, temperate needleleaved evergreen trees but negative for temperate and boreal broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{area}$ and leaf lifespan was positive globally but negative for evergreen shrubs and not significant for any other PFTs.
The correlation between leaf lifespan and $R_{d,area}$ significant and negative globally, but was not significant within any PFTs.
The only significant correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees.

A large number of pairwise trait correlations were not significant.
In some cases, this was driven by sample size (Fig.\ 1).
For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their area-normalized traits were significantly correlated with leaf lifespan.
In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ 5).
The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
Correlation strength was often, but not always, related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.

>>>>

For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.\ 4).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs, tropical broadleaved deciduous trees, temperate needleleaved evergreen trees but negative for temperate and boreal broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{area}$ and leaf lifespan was positive globally but negative for evergreen shrubs and not significant for any other PFTs.
The correlation between leaf lifespan and $R_{d,area}$ was significant and negative globally, but was not significant within any PFTs.
Finally, although we did not include maximum photosynthesis rate ($A_{max}$)---a key trait in the canonical leaf economic spectrum (Wright et al.\ 2004)---$V_{c,max,mass}$ and $J_{max,mass}$ generally exhibited the expected positive correlations with SLA and negative correlations with leaf lifespan, as did $R_{d,mass}$, though many correlations were not significant, primarily due to sample size.

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ 5).
For example, the correlations of SLA with $N_{mass}$ and $P_{mass}$ were generally weaker in needleleaved PFTs than in broadleaved PFTs.
On the other hand, relationships of SLA and $N_{area}$ were strongly negative for all PFTs (except the data-limited needleleaved deciduous trees), and temperate needleleaved species in particular had among the strongest SLA-$N_{area}$ and SLA-$P_{area}$ correlations of any PFT.

The strength of the correlations of $N$ with photosynthetic parameters $V_{c,max}$ and $J_{max}$ varied by plant functional type, regardless of whether the traits were normalized by mass or area.

(TODO Flesh out this text).

The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.

In some cases, this variability was driven by low sample sizes (Fig.\ 1, S1; Tab.\ S3, S4).
For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
Similarly, we had no observations of dark respirations for deciduous boreal shrubs, which explains why we found no significant correlations of dark respiration with any other trait for that PFT.
That being said, the relationship between correlations strength and sample size was very inconsistent (Fig.\ S1; Tab.\ S4).
For example, temperate needleleaf evergreen trees and C3 arctic grasses showed the strongest $N_{area}$-$P_{area}$ correlations despite having relatively average sample sizes for both traits.
In fact, every trait pair we investigated had at least one (and often several) cases where a PFT with a larger sample size showed weaker correlations than PFTs with lower sample sizes, or where correlation strength varied significantly among PFTs with similar sample sizes (Fig.\ S1).
Relationships between correlation strength and data range were even less consistent (Fig.\ S2).
Based on these results, we conclude that the variation we observed in our correlation results is not driven primarily by sampling artifacts and at least partially captures some underlying ecophysiological differences between PFTs.

### 2018-06-21 supplementary_information.Rmd
Add some correlation text to manuscript

correlation_scatter_cap <- paste(
  "Pairwise mean squared trait correlation values as a function of sample size.",
  "Note that the x-axis uses a $\\log_{10}$ scale."
)

@@

```{r mean_value_table, echo = FALSE, results = "asis"}
table_cap_mass <- paste(
  "Mean and 95\\% confidence interval of trait estimates",
  "for mass-normalized traits from the hierarchical model."
)
table_cap_area <- gsub("mass", "area", table_cap_mass)

@@

```{r sample_size_table, echo = FALSE, results = "asis"}
ssw <- readd(sample_size_wide)
kable(
  ssw,
  col.names = c(
    "PFT", "Leaf lifespan", "SLA",
    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
  ),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  group_rows("Broadleaf", 2, 6) %>%
  group_rows("Needleleaf", 7, 9) %>%
  group_rows("Shrub", 10, 12) %>%
  group_rows("Grass", 13, 15) %>%
  landscape()
```

>>>>

correlation_scatter_cap <- paste(
  "Relationship between squared pairwise correlation and sample size of pairwise observations for each PFT.",
  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values.",
  "Note that the x-axis uses a $\\log_{10}$ scale."
)

@@

```{r mean_value_table, echo = FALSE, results = "asis"}
table_cap_mass <- paste(
  "Mean and 95\\% confidence interval of trait estimates",
  "for mass-normalized traits from the hierarchical model.",
  "Note that some traits for some PFTs have effectively no constraint relative to the mostly uninformative prior (as evidenced by their very large confidence intervals),",
  "so their values should be used with caution."
)
table_cap_area <- gsub("mass", "area", table_cap_mass)

@@

```{r sample_size_table, echo = FALSE, results = "asis"}
ssw <- readd(sample_size_wide)
kable(
  ssw %>%
    dplyr::select(pft, leaf_lifespan, SLA,
                  Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
                  Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
  col.names = c(
    "PFT", "Leaf lifespan", "SLA",
    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
  ),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  group_rows("Broadleaf", 2, 6) %>%
  group_rows("Needleleaf", 7, 9) %>%
  group_rows("Shrub", 10, 12) %>%
  group_rows("Grass", 13, 15) %>%
  landscape()
```
